Weather covariance matrices as U_hyp for GR50

### 2021-01-12 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Move mash files to analysis/mash directory

## 3a. Covar Loadings
How many SNP effects load onto each covariance matrix?
```{r}
gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf.rds")
midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects33000SNPs_FL50_Midwest.rds")
gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/gwas/BLUPs/single_sites/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")
```

>>>>

## 3a. Covar Loadings
How many SNP effects load onto each covariance matrix?
```{r}
gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf.rds")
midwm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects33000SNPs_FL50_Midwest.rds")
gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_FL50_Gulf_and_Midwest.rds")
```

### 2021-01-12 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Move mash files to analysis/mash directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
```

@@

### run mash 
```{r}
for(k in c(1, 3)){

  numSNPs <- 19000
  m_out <- mash_standard_run(path = file.path(outputdir, names(subpop_v)[k]),
                             numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("GR50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
}

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
```

@@

### run mash 
```{r}
for(k in c(1,3,2)){

  numSNPs <- 19000
  m_out <- mash_standard_run(path = file.path(outputdir), #, names(subpop_v)[k]), # use to run on genegenie within mash dirs
                             numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("GR50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
}

@@

## 3a. Covar Loadings
How many SNP effects load onto each covariance matrix?
```{r}
gulfm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects19000SNPs_GR50_Gulf.rds")

gandm <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects19000SNPs_GR50_Gulf_and_Midwest.rds")
```
44765 SNPs tested for G&M, 45931 SNPs tested for Gulf.

Midwest I haven't run successfully yet as of 2021-01-12 - I will need to remove some locations with all-zero effects, probably.

### ii. Gulf
```{r}
length(get_significant_results(gulfm))  # 9142/45931  # 19.9%
get_significant_results(gulfm)[1:10]

covargulf <- mash_plot_covar(gulfm, suffix = "19000SNPs_GR50_Gulf", saveoutput = TRUE)
covargulf
get_U_by_mass(gulfm)
mash_plot_Ulist(gulfm, range = get_U_by_mass(gulfm), limits = FALSE)
mash_plot_manhattan_by_condition(gulfm, saveoutput = FALSE)
mash_plot_sig_by_condition(gulfm, saveoutput = FALSE)
```

### ii. G&M
```{r}
length(get_significant_results(gandm))  # 15104/44765  # 33.7%
get_significant_results(gandm)[1:10]

covargam <- mash_plot_covar(gandm, suffix = "19000SNPs_GR50_Gulf_and_Midwest", saveoutput = TRUE)
covargam
get_U_by_mass(gandm)
mash_plot_Ulist(gandm, range = get_U_by_mass(gandm), limits = FALSE)
mash_plot_manhattan_by_condition(gandm, saveoutput = FALSE)
mash_plot_sig_by_condition(gandm, saveoutput = FALSE)
```

```{r}
covargam$covar_df %>% 
  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("identity", `Covariance Matrix`) ~ "canonical",
                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
                          TRUE ~ "other"
                          )) %>%
  group_by(Type) %>%
  summarise(Mass = sum(Mass))
```
Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M

Not stunning, especially compared to how the flowering hyp do.

## 3b.	Mashhattans 
Mashhattans colored by covariance matrices.

### ii. Plot setup

Get posterior weight dataframe for the covariance matrix grid. Per SNP/row, entries in this grid sum to 1.

Get log10BayesFactor for mash SNPs also and combine these into a dataframe for ggplot.
```{r}
m <- gandm
log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
    as.data.frame() %>%
    rownames_to_column(var = "value") %>%
    mutate(value = as.integer(.data$value)) %>%
    as_tibble() %>%
    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
    dplyr::rename(log10BayesFactor = .data$V1) %>%
    dplyr::select(-.data$value)

#df <- get_topcov(m)

pw_df <- as_tibble(m$posterior_weights)
pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
  bind_cols(pw_df)

ggman_df <- pw_df %>%
  #left_join(m_meta) %>%
  separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
           extra = "merge", convert = TRUE) %>%
  left_join(log10bf_df, by = "Marker") %>%
  arrange(.data$Chr, .data$Pos)

log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))

mash_plot_covar(m)
```

Make a list containing all of the ggplots for all entries in the covariance grid represented by `pw_df`.
Then view all the ggplots in gglist.
```{r}
gglist <- list()
for(i in seq_along(names(pw_df)[-(1:2)])){
var <- names(pw_df)[i+2]   
gglist[[i]] <- ggman_df %>%
  arrange(!! sym(var)) %>%
  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
  switchgrassGWAS::theme_oeco +
  geom_point(aes(color = !! sym(var), fill = !! sym(var),
                 shape = as.factor(.data$Chr))) +
  facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
  scale_color_viridis_c(option = "D", end = 0.95) + 
  scale_fill_viridis_c(option = "D", end = 0.95) +
  theme(strip.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 8),
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
  labs(x = "Chromosome", y = log10BF) +
  scale_x_continuous(expand = c(0.32, 0.32)) +
  scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
  geom_hline(yintercept = 2, linetype = 2) +
  ggtitle(var)
}
gglist
```

Which covariance grids have SNPs with log10BF > 2 that load strongly onto them?
For Gulf mash:
GR50_Gulf.6: many, almost all < 2
GR50_Midwest.6: many, almost all < 2
equal_effects.7: many, almost all < 2
GR50_Gulf.6.7: many, almost all < 2
equal_effects.8: some, all <2
ED_PCA_1.9 many, many > 2
ED_PCA_2.9 many, few > 2
ED_PCA_4.9 some, few > 2
ED_tPCA.9: many, many > 2
ED_PCA_3.10: many, few > 2
ED_PCA_5.10: some, few > 2
OK.10 many, some > 2, many >0.8 loading & > 2
SD.10 many, some > 2
cgdd_12c_G&M.10: few, few > 2, some with 80%+ loading here

ED_tPCA.12: many, many > 2, many >0.8 loading esp at high log10BF
SD.12: many, many > 2, many > 0.8 loading esp at high log10BF

TX1.15: few bg, few >2

```{r}
var <- names(ggman_df)[ncol(ggman_df)-1]  
ggman_df %>%
  pivot_longer(cols = null:var, names_to = "covar_grid", values_to = "loading") %>%
  filter(loading > 0.5) %>%
  group_by(log10BayesFactor > 2, covar_grid, loading > 0.8) %>%
  tally()
```
## 3c. Data-driven mat

What patterns do the data-driven matrices capture that our weather matrices don’t capture?

SVD and plots of effects for the ED_PCA matrices.

## 3d. DS & AP

We characterized patterns of differential sensitivity and antagonistic pleiotropy between all SNPs and all pairs of gardens.

## 3e. Enrichments
We looked for enrichments of genes that could potentially affect flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering.

## 3f. Loglikelihood

log likelihood of models with and without U_hyp.
